---
title: ''
csl: peerj.csl
fontsize: 11pt
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
  word_document: default
geometry: margin=1.0in
bibliography: references.bib
---

```{r knitr_settings, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

source('../code/functions.R')

loadLibs(c("tidyverse", "scales", "knitr", "rmarkdown", "stringr", "FSA"))

```

```{r rdata_tables, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

#Read in the meta data file
metadata <- read_csv("../data/process/tables/meta_data.csv")

# Read in fecal count data
fecal_numOTU_data <- read_csv("../data/process/tables/fecal_sub_sample_1000_count_table.csv") %>% 
  mutate(taq = factor(taq, 
                      levels = c("ACC", "K", "PHU", "PL", "Q5"), 
                      labels = c("Accuprime", "Kappa", "Phusion", "Platinum", "Q5")), 
         cycles = factor(cycles, 
                            levels = c("15x", "20x", "25x", "30x", "35x"), 
                            labels = c("15x", "20x", "25x", "30x", "35x"))) %>% 
  group_by(taq, cycles) %>% 
  summarise(taq_median = median(numOTUs, na.rm = T), 
            taq_min = min(numOTUs, na.rm = T), 
            taq_max = max(numOTUs, na.rm = T)) %>% 
  mutate(range_diff = taq_max - taq_min)
  
# Read in fecal sample number of OTU results
fs_anova_summary <- as.data.frame(read_csv("../data/process/tables/fecal_overall_anova_results.csv"))
fs_tukey_summary <- as.data.frame(read_csv("../data/process/tables/fecal_overall_tukey_results.csv")) %>% 
  rename(padj = `p adj`)


# Read in the Mock count data
mock_num_otu_data <- read_csv("../data/process/tables/mock_sub_sample_1000_count_table.csv") %>% 
  mutate(taq = factor(taq, 
                      levels = c("ACC", "K", "PHU", "PL", "Q5"), 
                      labels = c("Accuprime", "Kappa", "Phusion", "Platinum", "Q5"))) %>% 
  group_by(taq, cycles) %>% 
  summarise(taq_median = median(numOTUs, na.rm = T), 
            taq_min = min(numOTUs, na.rm = T), 
            taq_max = max(numOTUs, na.rm = T)) %>% 
  mutate(range_diff = taq_max - taq_min)

ref_mock_otu_num <- read_tsv("../data/process/zymo_mock_only.shared") %>% 
  select(numOtus)


# Read in mock sample number of OTU results
mock_anova_summary <- as.data.frame(read_csv("../data/process/tables/mock_overall_anova_results.csv"))
mock_tukey_summary <- read_csv("../data/process/tables/mock_overall_tukey_results.csv")

# Read in error rate results
krsukal_error_rate <- read_csv("../data/process/tables/mock_error_overall_kruskal_results.csv")
dunn_error_rate <- read_csv("../data/process/tables/mock_error_overall_dunn_results.csv")

# Read in sequence error count results
kruskal_seq_error_count <- read_csv("../data/process/tables/mock_error_count_overall_kruskal_results.csv")
dunn_seq_error_count <- read_csv("../data/process/tables/mock_error_count_overall_dunn_results.csv")

# Read in chimera prevalence results
kruskal_chimera_prev <- as.data.frame(read_csv("../data/process/tables/mock_chimera_overall_kruskal_results.csv"))
dunn_chimera_prev <- read_csv("../data/process/tables/mock_chimera_overall_dunn_results.csv")

# Read in results from PERMANOVA and bray analysis
bray_sim_to_prev_cycle_kruskal_results <- as.data.frame(read_csv("../data/process/tables/bray_sim_to_prev_cycle_kruskal_results.csv"))
bray_permanova_by_taq_results <- as.data.frame(read_csv("../data/process/tables/bray_permanova_by_taq_results.csv"))

# Read in actual error data
error_data <- read_csv("../data/process/tables/mock_chimera_error_summary.csv") %>% 
                       filter(total_seqs >= 1000)

error_summary <- error_data %>% 
  group_by(taq, cycles) %>% 
  summarise_at(vars(mean_error, chimera_prevalence, seq_error_prevalence), funs(median, min, max)) %>% 
  select(taq, cycles, contains("chimera"), contains("mean_error"), contains("seq_error"))
  

# Vector of sub samples used
dist_data <- read_csv("../data/process/tables/bray_5_cycle_dist_1000_data.csv") %>% 
  mutate(taq = factor(taq, 
                      levels = c("ACC", "K", "PHU", "PL", "Q5"), 
                      labels = c("Accuprime", "Kappa", "Phusion", "Platinum", "Q5"))) %>% 
  group_by(taq, cycle_compare) %>% 
  summarise(dist_median = median(distance, na.rm = T), 
            dist_min = min(distance, na.rm = T), 
            dist_max = max(distance, na.rm = T)) %>% 
  mutate(range_diff = dist_max - dist_min) %>% 
  ungroup() %>% 
  filter(cycle_compare != "25to35")


fecal_kruskal_test_by_cycle_compare_summary <- dist_data %>% 
  nest() %>% 
  mutate(kruskal_analysis = map(data, ~kruskal.test(dist_median ~ factor(cycle_compare), data = .x)), 
         summary_data = map(kruskal_analysis, broom::tidy)) %>% 
  select(summary_data) %>% unnest(summary_data)
  
dunns_test_by_cycle_compare_summary <- dist_data %>% 
  nest() %>% 
  mutate(dunn_analysis = map(data, ~dunnTest(dist_median ~ factor(cycle_compare), data = .x, method = "bh")[["res"]])) %>% 
  select(dunn_analysis) %>% unnest() %>% 
  arrange(Comparison)

mock_dist_data <- read_csv("../data/process/tables/mock_bray_5_cycle_dist_1000_data.csv") %>% 
  mutate(taq = factor(taq, 
                      levels = c("ACC", "K", "PHU", "PL", "Q5"), 
                      labels = c("Accuprime", "Kappa", "Phusion", "Platinum", "Q5"))) %>% 
  filter(!is.na(distance)) %>% 
  group_by(taq, cycle_compare) %>% 
  summarise(dist_median = median(distance, na.rm = T), 
            dist_min = min(distance, na.rm = T), 
            dist_max = max(distance, na.rm = T)) %>% 
  mutate(range_diff = dist_max - dist_min) %>% 
  ungroup() %>% 
  filter(cycle_compare != "25to35")

mock_kruskal_test_by_cycle_compare_summary <- mock_dist_data %>% 
  nest() %>% 
  mutate(kruskal_analysis = map(data, ~kruskal.test(dist_median ~ factor(cycle_compare), data = .x)), 
         summary_data = map(kruskal_analysis, broom::tidy)) %>% 
  select(summary_data) %>% unnest(summary_data)

# Read in the regression data
r2_table <- read_csv("../data/process/tables/r2_otu_vs_chimera_summary.csv")

# Read in GC data
gc_data <- read_csv("../data/process/tables/v4_gc_wilcox_test_summary.csv")

# Read in RF model data
dunn_rf_data <- read_csv("../data/process/tables/dunn_rf_model_test_summary.csv")

```



# Assessing the Differences in 16S rRNA Gene Sequencing Due to High Fidelity DNA Polymerase 

\begin{center}
\vspace{25mm}

Marc A Sze${^1}$ and Patrick D Schloss${^1}$${^\dagger}$

\vspace{20mm}

$\dagger$ To whom correspondence should be addressed: pschloss@umich.edu

$1$ Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI




\end{center}

Co-author e-mails:

* marcsze@med.umich.edu



\newpage
\linenumbers

## Abstract

**Background.** A typical 16S rRNA gene sequencing workflow can be divided into preservation, extraction, amplification, and sequencing steps. At each of these stages error can be introduced that will change the underlying bacterial community composition results. In this study we focus on the amplification step's contribution to this overall error. To accomplish this we assessed 16S rRNA gene sequencing results in human fecal and mock community samples after using different high fidelity (HiFi) DNA polymerases and number of amplification cycles.  

**Methods.** We extracted DNA from fecal samples (n=4) using a PowerMag DNA extraction kit with a 10 minute bead beating step and amplified at 15, 20, 25, 30, and 35 cycles using Accuprime, Kappa, Phusion, Platinum, or Q5 HiFi DNA polymerase. Amplification of mock communities (technical replicates n=4) consisting of previously isolated whole genomes of 8 different bacteria used the same approach. We first assessed GC dependent differences, error rate, sequence error prevalence, chimera prevalence, and correlation between chimera prevalence and number of Operational Taxonomic Units (OTUs) by polymerase and number of cycles. Next, differences in the number of OTUs and taxa was examined based on the polymerase and number of cycles used. Additionally, differences in the bacterial community composition by the Bray-Curtis index also was assessed based on polymerase and number of cycles. Finally, Random Forest models were created to test whether the bacterial community was better at classifying polymerases, number of cycles, or individual donor. We also assessed whether the most important taxa in the polymerase and number of cycle Random Forest models also were the most important in the model for individual donors.    

**Results.** Although the total counts for specific taxa varied by polymerase, we found no noticeable differences in total counts based on high and low GC content (P-value $\geqslant$ `r round(min(pull(gc_data, p.value)), digits = 2)`). Chimera prevalence in mock communities varied by polymerase with differences being most notable at 35 cycles (Kappa = `r round((filter(error_summary, cycles == "35x", taq == "K") %>% pull(chimera_prevalence_median))*100, digits = 2)`% (median) versus Platinum = `r round((filter(error_summary, cycles == "35x", taq == "PL") %>% pull(chimera_prevalence_median))*100, digits = 2)`%) and this variation persisted after chimera removal using VSEARCH. We also observed positive correlations between chimera prevalence and the number of OTUs with Platinum having the highest (R^2^ = `r max(pull(r2_table, r2))`) and Kappa having the worst (R^2^ = `r min(pull(r2_table, r2))`). When analyzing mock community samples the variation in the number of OTUs detected by the polymerases was observable as early as 20 cycles (P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`). There also was a large range in the number of OTUs amplified by the polymerases at 35 cycles (Accuprime = `r filter(mock_num_otu_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_min)` - `r filter(mock_num_otu_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_max)` versus Phusion = `r filter(mock_num_otu_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(mock_num_otu_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`). When analyzing fecal samples we observed that the range in the number of OTUs detected was not consistent between HiFi DNA polymerases (eg. at 35 cycles Accuprime = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_max)` (min - max) versus Phusion = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`). Additionally, the median number of OTUs varied by HiFi DNA polymerase used at 35 cycles (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "35x")[, "bh"]) < 0.0001, invisible("<"), invisible("check data"))` 0.0001). Random Forest models were most successful at classifying individual donor samples rather than polymerase or number of cycles used (P-value $\leqslant$ `r format(max(filter(dunn_rf_data, str_detect(var1, "fecal"), str_detect(var2, "fecal")) %>% pull(bh)), scientific = T, digits = 3)`). Additionally, the most important OTUs in the polymerase and number of cycle models were not the most important in the individual donor sample model.

**Conclusions.** Although there are 16S rRNA gene sequencing differences based on polymerase and number of cycles used, they are small with respect to the biological differences between indivdiuals. Additionally, our observations were not able to identify GC-based differences in amplification based on the number of cycles used. Collectively, these results provide evidence that a real biological difference between groups, based on 16S rRNA gene sequencing, should be detectable regardless of polymerase and number of cycles used.  



\newpage


## Introduction

The bacterial community is reported to vary between case and control for a number of diseases [@Turnbaugh2008; @Baxter2016; @Sze2015; @Bonfili2017]. However, for diseases like obesity, the taxa identified have varied widely depending on the study [@Turnbaugh2008; @Zupancic2012]. Some of this variation could be due to error introduced during the 16S rRNA gene sequencing workflow. Yet, standardizing a 16S rRNA gene sequencing workflow will ultimately result in a standardized and reproducible bias due to choices made on the methods used for preservation, extraction, PCR, and sequencing. Within this context, all 16S rRNA gene sequencing methods are biased even when these workflows are standardized to increase reproducibility. In order to interpret specific studies within the broader context of the overall field, assessing error at different parts of the 16S rRNA gene sequencing workflow is critical.

A typical 16S rRNA gene sequencing workflow can be divided into preservation, extraction, PCR, and sequencing steps. The preservation and extraction stages of the 16S rRNA gene sequencing workflow have been the most extensively studied [@contamination_Salter2014; @preservation_Song_2016; @storage_Bassis_2017; @review_Kim_2017]. For preservation and extraction stages of the workflow, it has been consistently found that there are biases based on the kits used, but that these differences are smaller than the overall biological difference measured between samples with different kits [@preservation_Song_2016; @storage_Bassis_2017]. Since these studies use the same PCR approach while varying preservation or extraction method, the contribution of PCR bias to this overall workflow is not well characterized.

There is a large body of literature that shows there are biases due to primer and number of cycles chosen for the PCR stage of 16S rRNA gene sequencing [@Eckert1991; @Burkardt2000]. Primers have variable region dependent binding affinities which causes an inability to detect specific bacteria (e.g. V1-V3 does not detect *Haemophilus influenzae* and V3-V5 does not detect *Propionibacterium acnes*) [@Meisel2016; @Sze2015 (Table S4)]. Another source of error is the selective amplification of AT-rich over GC-rich sequences which exaggerate the difference between 16S rRNA genes higher in AT versus those higher in GC content [@polz_bias_1998]. Many of these sources of biases are made worse as the number of cycles increases [@Wang1996; @Haas2011; @Kebschull2015]. For example, both amplification error and non-specific amplification (e.g. incorrect amplicon size products) also can increase as the number of cycles increases. This will increase the number of Operational Taxonomic Units (OTUs) observed and drastically change the values obtained from commonly used diversity measures [@BautistadelosSantos2016; @Acinas2005]. Additionally, as the number of cycles increases more chimeras can form from an aborted extension step that causes a priming error and subsequent secondary extension [@Haas2011]. These chimeras will artificially increase community diversity by increasing the number of OTUs that are observed [@Haas2011]. In addition to these sources of errors, there also are multiple families of DNA polymerases that have their own error rate and proof reading capacity [@polymerase_Ishino_2014]. Interestingly, the influence that these different DNA polymerases can have on the observed 16S rRNA gene sequencing results have not been well studied like some of the other sources of PCR-based bias.     

A recent study found differences in the number of OTUS and chimeras between normal and high fidelity DNA polymerases [@taq_Gohl_2016]. The authors could reduce the difference between the two polymerases by optimizing the annealing and extension steps within the PCR protocol [@taq_Gohl_2016]. Yet, within this study there was no comparison made between different high fidelity DNA polymerases. Due to this gap, it is natural to extend this line of inquiry and test if biases in the number of OTUs and chimeras also are dependent on the type of high fidelity DNA polymerase. This study will investigate how high fidelity DNA polymerases can bias observed bacterial community results derived from 16S rRNA gene sequencing. We will accomplish this by examining the number of OTUs, Bray-Curtis index, error rate, number of sequences with an error, and chimera prevalence at varying number of cycles in five different high fidelity DNA polymerases



\newpage

## Materials & Methods

***Human and mock samples:***
Fecal samples were obtained from 4 individuals who were part of the Enterics Research Investigational Network (ERIN). The processing and storage of these samples were previously published [@erin_seekatz_2016]. Other than confirmation that none of these individuals had a *Clostridium difficle* infection, clinical data and other types of meta data were not utilized or accessed for this study. All samples were extracted using the MOBIO^TM^ PowerMag Microbiome RNA/DNA extraction kit (now Qiagen, MD, USA). The ZymoBIOMICS^TM^ Microbial Community DNA Standard (Zymo, CA, USA) was used for mock communities and was made up of *Pseudomonas aeruginosa*, *Escherichia coli*, *Salmonella enterica*, *Lactobacillus fermentum*, *Enterococcus faecalis*, *Staphylococcus aureus*, *Listeria monocytogenes*, and *Bacillus subtilis* at equal genomic DNA abundance (http://www.zymoresearch.com/microbiomics/microbial-standards/zymobiomics-microbial-community-standards).

***PCR protocol:***
The five different high fidelity DNA polymerases (hereto referred to as polymerases) that were tested included AccuPrime^TM^ (ThermoFisher, MA, USA), KAPA HIFI (Roche, IN, USA), Phusion (ThermoFisher, MA, USA), Platinum (ThermoFisher, MA, USA), and Q5 (New England Biolabs, MA, USA). The polymerases activation time was 2 minutes, unless a different activation was specified by the manufacturer. The annealing and extension time for Platinum and Accuprime followed a previously published protocol [@protocol_Kozich_2013] (https://github.com/SchlossLab/MiSeq_WetLab_SOP/blob/master/MiSeq_WetLab_SOP_v4.md). For Kappa and Q5, the annealing and extension time also were from a previously published protocol [@taq_Gohl_2016]. For Phusion, the company defined activation and annealing times were used while the extension time followed the Accuprime and Platinum settings. 

The number of cycles in the PCR for fecal and mock samples started at 15 and increased by 5 up to 35 cycles, with amplicons used at each 5-step increase for sequencing. The PCR of fecal DNA samples consisted of all 4 samples at 15, 20, 25, 30, and 35 cycles for each polymerase (total sample n=100). The mock communities had 4 replicates at 15, 20, 25, and 35 cycles and 10 replicates for 30 cycles for all polymerases (total samples n=130). No mock community sample had enough PCR product at 15 cycles for adequate 16S rRNA gene sequencing.    

***Sequence processing:***
The mothur software program was used for all sequence processing steps [@mothur_schloss_2009]. The protocol has been previously published [@protocol_Kozich_2013] (https://www.mothur.org/wiki/MiSeq_SOP). Two major differences from the published protocol were the use of VSEARCH instead of UCHIME for chimera detection and the use of the OptiClust algorithm instead of average neighbor for OTU generation at 97% similarity [@vsearch_Rognes_2016; @uchime_Edgar_2011; @opticlust_Westcott_2017]. Sequence error was determined using the 'seq.error' command on mock samples to compare back to the reference 16S sequences of *P. aeruginosa*, *E. coli*, *S. enterica*, *L. fermentum*, *E. faecalis*, *S. aureus*, *L. monocytogenes*, and *B. subtilis* [@mothur_schloss_2009; @vsearch_Rognes_2016; @rdp_Cole_2013].  

***Analysis workflow:***
To adjust for unequal sequencing, all samples were rarefied to 1000 sequences for downstream analysis. Analysis of the mock community of each polymerase for GC-based amplification differences, sequence error rate, number of sequences with an error, base substitution, and numbers of chimeras before and after chimera removal with VSEARCH was assessed. Additionally, the correlation between the number of chimeras and the number of OTUs was also assessed. The total number of OTUs, taxa differences, and Bray-Curtis indices were analyzed for both the fecal and mock community samples. Finally, Random Forest models were created to assess whether classification of polymerase, cycles, or individual could be performed best using 16S rRNA gene sequencing data. Additionally, overlap between the most important OTUs to the three models was assessed using mean decrease in accuracy (MDA).  

***Statistical analysis:***
All analysis was done with the R (v `r paste(version$major, ".", version$minor, sep = "")`) software package [@r_citation_2017]. Data transformation and graphing was completed using the tidyverse package (v `r paste(packageVersion("tidyverse")[1])`) and colors selected using the viridis package (v `r paste(packageVersion("viridis")[1])`) [@tidyverse_2017; @viridis_citation_2017]. Differences in the total number of OTUs were analyzed using an ANOVA with a tukey post-hoc test. For the comparison of the number of OTUS in fecal samples the data was normalized to each individual by cycle number to account for the biological variation. Bray-Curtis distance matrices were generated using mothur. The distance matrix data was analyzed using PERMANOVA with the vegan package (v `r paste(packageVersion("vegan")[1])`) [@vegan_citation] and Kruskal-Wallis tests within R. The Random Forest models were run using the caret package (v `r paste(packageVersion("caret")[1])`) [@caret_citation]. A total of 100 10-fold CV runs on different 80/20 splits of the data was run to generate a range of the Logloss value. The probability of a correct call was obtained from this Logloss value by taking the negative natural logrithim. For both error and chimera analysis, samples were tested using Kruskal-Wallis with a Dunns post-hoc test. Where applicable, correction for multiple comparison utilized the Benjamini-Hochberg method [@benjamini_controlling_1995].    

***Reproducible methods:***
The code and analysis can be found here https://github.com/SchlossLab/Sze_PCRSeqEffects_XXXX_2017. The raw sequences can be found on the SRA at the following accession number SRP132931.


\newpage


## Results

***The number of OTUs generated are dependent on polymerase used:***
Differences in the range of the number of OTUs detected for fecal samples is dependent on the polymerase used (e.g. Accuprime at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_max)` versus Phusion at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`) [Figure 1]. Additionally, there is a trend for lower number of cycles (15-20) to result in a reduced range in the number of OTUs versus higher number of cycles (25, 30, and 35) for all polymerases (e.g. Phusion at 15 cycles = `r filter(fecal_numOTU_data, cycles == "15x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "15x", taq == "Phusion") %>% pull(taq_max)` versus Phusion at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`) [Figure 1]. There is an overall difference in the number of OTUs detected within fecal samples between polymerases at 35 cycles (F-stat > `r round(min(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "F.value"]), digits = 2)`, P-value = `r format(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "bh"], digits = 2, nsmall = 2, scientific = T)`) [Table S1], but a Tukey post-hoc test failed to identify which specific polymerase groups were different (P-value `r ifelse(min(filter(fs_tukey_summary, sub_sample_level == 1000)[, "padj"]) > 0.05,  invisible(">"), invisible("check data"))` 0.05) [Table S2]. The polymerase dependent difference in the range of the number of OTUs also was observed in the mock community samples [Figure 2]. The closest the polymerases came to the total of `r pull(ref_mock_otu_num, numOtus)` OTUs created by the mock reference 16S sequences was at 25 and 30 cycles [Figure 2]. Regardless of if fecal or mock communities were used, the same polymerases generated high and low number of OTUs and this was consistent across the number of cycles used [Figure 1-2 & Table S1-S2].  In contrast to the results obtained with fecal samples, differences between polymerases for the number of OTUs created were observed as early as 20 cycles in the mock community (F-stat = `r round(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "F.value"], digits = 2)`, P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`) [Table S1]. Using a Tukey post-hoc test, the majority of differences for the number of OTUs detected in the mock community was largely due to Kappa and Platinum versus the other polymerases across the number of cycles used [Table S2]. Based on these observations in fecal and mock communities, it is clear that using different polymerases result in a different total number of OTUs within a sample. 


***The bacterial communities generated by the polymerases were similar but varied by number of cycles:***
Within each respective 5-cycle increment comparison there was no difference in Bray-Curtis index between the polymerases (P-value `r ifelse(min(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000)[, "bh"], na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05) [Figure 3]. Using PERMANOVA to test for community differences based on any of the number of cycles within polymerases, only Phusion had cycle dependent differences (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`). For fecal samples, Phusion was one of two polymerases that had enough sequences to be rarefied to 1000 sequences at 15 cycles. There was an overall decrease in Bray-Curtis index by cycle comparison group (e.g. the 15 cycle versus 20 cycle group compared to the 30 cycle versus 35 cycle group) (P-value `r ifelse(pull(fecal_kruskal_test_by_cycle_compare_summary, p.value) < 0.01, invisible("<"), invisible(">"))` 0.01) [Figure 3A]. Using a Dunn's post-hoc test the 15 cycle versus 20 cycle and 20 cycle versus 25 cycle comparison groups had a higher Bray-Curtis index than the 30 cycle versus 35 cycle comparison group (P-value `r ifelse(min(pull(dunns_test_by_cycle_compare_summary, P.adj)) < 0.05, invisible("<"), invisible(">"))` 0.05). Similar trends were observed in the mock community but none were significant (P-value `r ifelse(pull(mock_kruskal_test_by_cycle_compare_summary, p.value) < 0.05, invisible("<"), invisible(">"))` 0.05) [Figure 3B]. Overall, these data suggest that the number of cycles can change the bacterial community independent of the polymerase used to generate the sequences.


***Sequence error varies by polymerase and is consistent across the number of cycles used:***
The median error rate varied by polymerase, with Kappa having the highest error rate of all the polymerases across the number of cycles [Figure 4 & Table S3]. The majority of the differences across the number of cycles was between Kappa and the other polymerases [Figure 4 & Table S4]. The total sequences with at least one error was also polymerase dependent, with the majority of differences being between Kappa or Accuprime and the other polymerases [Table S3 & S4]. These differences in error rates were not due to polymerase dependent differences in base substitution rate [Figure S1]. Collectively, the results suggest that sequence error is dependent on polymerase, persists across the number of cycles used, and are not due to any bias towards a specific base substitution.



***Prevalence of chimeric sequences are polymerase dependent and correlate with the number of OTUs:***  Based on the previous results, we examined whether chimeras also were dependent on polymerase and whether this could affect the number of OTUs. We observed significant differences in the chimera prevalence based on polymerase across the number of cycles used (P-value `r ifelse(max(filter(kruskal_chimera_prev)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S3]. Differences in chimera prevalence between Platinum and all other polymerases accounted for the majority of these differences [Table S4]. Accuprime^TM^ had the lowest chimera prevalence of all polymerases regardless of whether chimera removal with VSEARCH was used [Figure 5A & 5B]. The number of cycles used clearly increased chimera prevalence for all polymerases but the rate of increase differed [Figure 5A & 5B]. Additionally, there was a plateau in the total percent of chimeras that were removed that was similar for all polymerases [Figure 5C]. A positive correlation was observed between chimeric sequences and the number of OTUs for all polymerases [Figure 6]. This positive correlation was strongest for Accuprime^TM^, Platinum, and Phusion [Figure 6]. This data suggests that chimera prevalence depends on polymerase, is made worse by increasing the number of cycles, and confirms that the number of OTUs is dependent on the prevalence of these chimeric sequences. 

\newpage

## Discussion

Our observations support that the number of OTUs, error rate, and chimera prevalence depends on polymerase used and that chimera prevalence is highly correlated with the number of OTUs [Figure 1-2 & 4-6]. Our observations also suggest that undetected chimeras, not removed using standard bioinformatic approaches, cause many of these differences. In addition to the reported polymerase dependent differences, the bacterial community also varied by the number of cycles used. These differences are important because many diversity metrics rely on calculations (e.g. richness) that are dependent on the number of OTUs. Since the number of OTUs are dependent on error rate and chimera prevalence the diversity metrics can vary simply by changing the polymerase used. Based on these observations, metrics that measure diversity using calculations such as richness depend on the polymerase, but this may not be the case for all diversity metrics.    

Based on polymerase used, our observations generally found no difference when using the Bray-Curtis index to measure the bacterial community. Specifically, there was no difference in distance between successive 5-cycle increments within samples between the polymerases when using the Bray-Curtis index [Figure 3]. However, the distance between cycle comparison groups (e.g. 20 cycles versus 25 cycles group compared to 30 cycles versus 35 cycles group) was different across polymerase used. These results support the idea that cycle number and not the specific polymerase used may have a larger affect on metrics such as Bray-Curtis. One possible reason for this outcome is that many of the OTUs generated by polymerase dependent error rates and chimera prevalence are likely not highly abundant, allowing the Bray-Curtis index to be able to successfully down-weight these OTUs [@bc_index_Minchin1987]. The choice of downstream diversity metric could be an important consideration in helping to mitigate some of the observed polymerase dependent differences in error rate and chimera prevalence. Metrics that solely use presence/absence of OTUs (e.g. Jaccard [@Real1996], richness) may be less robust to polymerase dependent error rates and chimera prevalence. When choosing a distance metric, careful consideration of the biases introduced from the PCR step of the 16S rRNA gene sequencing workflow need to be taken into account. One possible way to better choose polymerases may be based on the DNA polymerase family used in the PCR mixture.

Although the variation in error rate and chimera prevalence may be due to the DNA polymerase family because of the different binding affinities and error correction capacity [@polymerase_Ishino_2014], this is unlikely to be the only contributor. Within our study the highest and lowest chimera prevalence both belonged to a family A polymerase (Platinum and Accuprime^TM^ respectively) [@polymerase_Ishino_2014]. Additionally, based on the information supplied by the respective manufacturers, the differences between the two PCR mixtures are not immediately apparent. Both PCR mixtures contain a recombinant *Taq* DNA polymerase, a *Pyrococcus* spp GB-D polymerase and a platinum *Taq* antibody. Since it is not possible to know everything about the mixture beyond what was willingly provided by the manufacturer, it is possible that differences in how the recombinant *Taq* was generated or other compounds within the PCR mixture could be a contributing factor for the differences in error rate and chimera prevalence. Beyond the choice of the type of polymerase, there may be other ways to reduce the affect of polymerase dependent error rates and chimera prevalence on the downstream results. 

Standardization of the 16S rRNA gene sequencing workflow may partially solve this problem by introducing a consistent bias to all samples. However, as mentioned in the introduction, many of the choices made to standardize the workflow will result in missing important members of the microbial community. Therefore, the degree of workflow standardization across studies and research groups need to be approached on a study by study basis. All aspects of the 16S rRNA gene sequencing workflow should be customized for the specific scientific questions that want to be answered. Although a diversity of approaches may make reproducibility and replicability more difficult it will help reduce systematic biases from occurring. 





```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Results have an impact on downstream diversity measurements due to effects on total OTUs


#Need to cover that different types of polymerase may affect outcome
#- Accuprime HiFi from recombinant A-Family Taq DNA polymerase, Pyrococcus spp GB-D polymerase (proof-reading), platinum Taq Antibody (inhibits room temp activity)

#- Kappa HiFi from B-family DNA polymerase (proof-reading baked in), propietary antibody (inhibit room temp activity)

#- Q5 HiFi from B-family thermostable DNA polymerase with 3´→ 5´ exonuclease activity, fused to a processivity-enhancing Sso7d domain to support robust DNA amplification

#-Phusion HiFi B-family DNA polymerase a novel Pyrococcus-like enzyme fused with a processivity-enhancing domain, generate blunt-ended products.

#- Platinum HiFi recombinant A-Family Taq DNA polymerase, Pyrococcus species GB-D polymeras, Platinum Taq Antibody

```



\newpage

## Conclusion

Our observations show that errors dependent on either high fidelity DNA polymerases used or the number of cycles chosen can change diversity metrics used in 16S rRNA gene sequencing. Care should be taken when choosing a polymerase for 16S rRNA gene surveys because they can influence diversity-based metrics. By knowing the inherent bias associated with different polymerases it allows for more informed choices to be made on how to reduce 16S rRNA gene sequencing-based error.

\newpage



## Acknowledgements

The authors would like to thank all the study participants in ERIN whose samples were utilized. We also would like to thank Judy Opp and April Cockburn for their effort in sequencing the samples as part of the Microbiome Core Facility at the University of Michigan. Additional thanks to members of the Schloss lab and Dr. Marcy Balunas for reading earlier drafts of the manuscript and providing helpful critiques. Salary support for Marc A. Sze came from the Canadian Institute of Health Research and NIH grant UL1TR002240. Salary support for Patrick D. Schloss came from NIH grants P30DK034933 and 1R01CA215574.   

\newpage

## References

<div id="refs"></div>

\newpage


**Figure 1: The number and range of OTUs in fecal samples varies by polymerase.** The points represent the median number of OTUs of all four fecal samples. The lines represent the range of the minimum and maximum number of OTUs detected within the four fecal samples. The range in the number of OTUs detected in the different fecal samples increased as cycle number increased. This increased range also was larger for specific HiFi DNA polymerases. 

**Figure 2: The number and range of OTUs in mock samples varies by polymerase.** The points represent the median number of OTUs for the mock samples. The lines represent the range of the minimum and maximum number of OTUs detected within the four fecal samples. The dotted black line represents the number of OTUs detected when only the references sequences for the mock community are clustered. The range in the number of OTUs detected in the mock samples increased as cycle number increased. This range was also larger for specific HiFi DNA polymerases.

**Figure 3: Lower five-cycle intervals have larger differences than higher five-cycle intervals using Bray-Curtis.** A) within person differences based on the next 5-cycle PCR interval in fecal samples. B) Within replicate difference based on the next 5-cycle PCR interval in mock samples. The points represent the median Bray-Curtis index for the samples. The lines represent the range of the minimum and maximum Bray-Curtis index value for each PCR 5-cycle increment comparison. The closer a sample is to a Bray-Curtis index of 1.00 the more dissimilar the bacterial community is of the two compared number of cycles. 

**Figure 4: Variation in error rate by polymerase in mock community samples are similar across the number of cycles used.** The error bars represent the 75% interquartile range of the median error rate.  

**Figure 5: Chimera prevalence varies by polymerase and increases with higher number of cycles used.** A) Percentage of chimeric sequences without the removal of chimeras with VSEARCH. C) Percentage of chimeric sequences with the removal of chimeras with VSEARCH. C) The total percent of chimeric sequences removed with VSEARCH by cycle number. The error bars represent the 75% interquartile range of the median. 

**Figure 6: Regardless of polymerase used the higher the chimera prevalence the higher the observed number of OTUs in mock community samples.** 

\newpage


**Figure S1: No preference for specific base subsitutions was observed across the polymerases in mock community samples.**  


