---
title: ''
csl: peerj.csl
fontsize: 12pt
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
  word_document: default
geometry: margin=1.0in
bibliography: references.bib
---

```{r knitr_settings, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

source('../code/functions.R')

loadLibs(c("tidyverse", "scales", "knitr", "rmarkdown", "stringr"))

```

```{r rdata_tables, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}


```



# What the Taq? The Influence of Taq Polymerase on 16S rRNA Sequencing

\begin{center}
\vspace{25mm}

Marc A Sze${^1}$ and Patrick D Schloss${^1}$${^\dagger}$

\vspace{20mm}

$\dagger$ To whom correspondence should be addressed: pschloss@umich.edu

$1$ Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI




\end{center}

Co-author e-mails:

* marcsze@med.umich.edu



\newpage
\linenumbers

## Abstract

**Background.** 

**Methods.**

**Results.** 

**Conclusions.** 




\newpage


## Introduction




\newpage

## Materials & Methods

***Insert Sub section here***



***Reproducible Methods:***
The code and analysis can be found here https://github.com/SchlossLab/Sze_PCRSeqEffects_XXXX_2017. 


\newpage


## Results

***Insert Sub title here***




\newpage

## Discussion



\newpage

## Conclusion



\newpage



## Acknowledgements

The authors would like to thank all the study participants who were a part of each of the individual studies utilized. We would also like to thank each of the study authors for making their data available for use. Finally we would like to thank the members of the Schloss lab for valuable feed back and proof reading during the formulation of this manuscript.

\newpage

## References

<div id="refs"></div>

\newpage


**Table 1: **

\footnotesize

```{r table1, echo=FALSE, message=FALSE, warning=FALSE}


```

\normalsize
\newpage

**Table 2: **

\footnotesize

```{r table2, echo=FALSE, message=FALSE, warning=FALSE}


```

\normalsize
\newpage



**Figure 1: .** 

**Figure 2: .** 

**Figure 3: .**

**Figure 4: .** 

**Figure 5: .** 

**Figure 6: .** 

**Figure 7: .**  


\newpage


**Figure S1: .**

**Figure S2: .** 

**Figure S3: .** 

**Figure S4: .** 

**Figure S5: .** 

**Figure S6: .** 

**Figure S7: .** 

\newpage
