---
title: ''
csl: peerj.csl
fontsize: 12pt
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
geometry: margin=1.0in
bibliography: null
---

# What the Taq? The Influence of Different Hi-Fidelity Taq Polymerase on 16S rRNA Sequencing

\vspace{10mm}

\begin{center}
Marc A Sze and Patrick D Schloss

\vspace{10mm}

Supplemental
\end{center}


\newpage

```{r knitr_settings, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

source('../code/functions.R')

loadLibs(c("tidyverse", "scales", "knitr", "rmarkdown"))

```


**Table S1: ANOVA Results of Number of OTUs Differences between HiFi Taq in Fecal Samples**

```{r supplemental_table1, echo=FALSE, message=FALSE, warning=FALSE}

fs_anova <- as.data.frame(read_csv("../data/process/tables/fecal_overall_anova_results.csv")) %>% 
  mutate_at(vars(Sum.Sq:F.value), funs(round(., digits = 2))) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2)))

colnames(fs_anova) <- c("DF", "Sum Squares", "Mean Squares", "F value", "P-value", "BH", "Cycle", "Sub-Sample Depth")

kable(fs_anova, format = "markdown", align = 'c', padding = 2)
```


\newpage


**Table S2: Tukey Post-Hoc Results of Number of OTUs Differences between HiFi Taq in Fecal Samples**

```{r supplemental_table2, echo=FALSE, message=FALSE, warning=FALSE}

fs_tukey <- as.data.frame(read_csv("../data/process/tables/fecal_overall_tukey_results.csv")) %>% 
  mutate_at(vars(diff:upr), funs(round(., digits = 2))) %>% 
  rename(padj = `p adj`) %>% 
  mutate(padj = as.numeric(padj)) %>% 
  filter(padj < 0.05) %>% 
  mutate_at(vars(padj), funs(format(., scientific = TRUE, digits = 2, nsmall = 2)))

colnames(fs_tukey) <- c("Difference", "Lower", "Upper", "P Adjusted", "Comparison", "Cycle", "Sub-Sample Depth")

kable(fs_tukey, format = "markdown", align = 'c', padding = 2)

```


\newpage


**Table S3: ANOVA Results of Number of OTUs Differences between HiFi Taq in Mock Samples**

```{r supplemental_table3, echo=FALSE, message=FALSE, warning=FALSE}

mock_anova <- as.data.frame(read_csv("../data/process/tables/mock_overall_anova_results.csv")) %>% 
  mutate_at(vars(Sum.Sq:F.value), funs(round(., digits = 2))) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2))) %>% 
  filter(sub_sample_level > 500)

colnames(mock_anova) <- c("DF", "Sum Squares", "Mean Squares", "F value", "P-value", "BH", "Cycle", "Sub-Sample Depth")

kable(mock_anova, format = "markdown", align = 'c', padding = 2)
```


\newpage


**Table S4: **

```{r supplemental_table4, echo=FALSE, message=FALSE, warning=FALSE}


```


\newpage

**Table S5: **

```{r supplemental_table5, echo=FALSE, message=FALSE, warning=FALSE}


```



\newpage


**Table S6: **

```{r supplemental_table6, echo=FALSE, message=FALSE, warning=FALSE}


```

\newpage

**Table S7: **

```{r supplemental_table7, echo=FALSE, message=FALSE, warning=FALSE}




```











